We are not the 99 percent: quantifying asphericity in the distribution of Local Group satellites

### 2018-02-23 paper/paper.tex
edit

\begin{table*}
\centering
\begin{tabular}{|p{4.0cm}|p{4.5cm}| p{5.5cm}| c|}\hline
Reference & Target Measurement & Parent Simulation & Probability ($\%$)\\\hline
\text{\cite{2006AJ....131.1405K}} & RMS height in 15 brightest
satellites & Monte Carlo satellite distributions with a power law
radial distribution & $87-99$\\
\text{\cite{2007MNRAS.374.1125M}} & Triaxiality and RMS height in 16 brightest satellites & Monte Carlo from an spherical power law radial
distribution & $17$\\
\text{\cite{2013ApJ...766..120C}}& RMS height in the 27 brightest
satellites & Monte Carlo randomized satellite distribution & High\\
This Work & Triaxiality and RMS height in 11-15 brightest
satellites & 24 halos from selected pairs in a cosmological N-body Dark Matter only simulation ($\sim
10^{6}$\Msun particle mass resolution)& 56 \\
This Work & Triaxiality and RMS height in 11-15 brightest
satellites & 20 halos from selected pairs in a cosmological N-body hydro simulation ($\sim
10^{6}$\Msun particle mass resolution)& 27 \\
This Work & Triaxiality and RMS height in 11-15 brightest
satellites & 12 high resolution DM only N-body simulations ($\sim
10^{5}$\Msun particle mass resolution) of halo pairs & 37 \\
\hline
\end{tabular}
\caption{Probability results derived in previous publications for the
  triaxiality and/or root mean squared (RMS) height of M31 satellites.
\label{table:M31}}
\end{table*}

@@

To summarize, we start by using asphericity to characterize in the
same way simulations and observations. 
Then, we build an explicit probability distribution for the asphericity and
use simulations to estimate the parameters in this
distribution. 
Finally, we use these distributions to generate large number of
samples and directly estimate the number of systems meeting a desired
set of criteria.

@@

\subsection{Describing Samples with the Inertia Tensor}
We base all our results on the description provided by the inertia
tensor defined by the satellites' positions.

@@

From this tensor we compute its eigenvalues,
$\lambda_1>\lambda_2>\lambda_3$, and corresponding eigenvectors,
$\hat{I}_1$, $\hat{I}_2$, $\hat{I}_3$.
We define the size of the three ellipsoidal axis as
$a=\lambda_1$, $b=\lambda_2$ and $c=\lambda_3$.
We also define $\hat{n}\equiv \hat{I}_1$ as the vector perpendicular to the
planar satellite distribution. 
We also define the width, $w$, of the planar satellite distribution,
as the standard deviation of all satellite distances to the plane
defined by the vector $\hat{n}$.

To summarize we characterize the satellite distribution by for
quantities obtained from the inertia tensor: 
\begin{itemize}
\item Plane width, $w$.
\item $c/a$ axis ratio.
\item $b/a$ axis ratio.
\end{itemize}

@@

We compare every satellite distribution against its own spherically
randomized distribution.
We keep fixed the radial position of every satellite
with respect to the central galaxy and then randomize its angular
position. 
We repeat this procedure 1000 times for each satellite distribution
and proceed to measure the quantities mentioned in the previous section:
$w$, $c/a$ and $b/a$.
For each quantity we compute the average and standard deviation from
the 1000 random samples. 
This allows us to build a normalized version of all quantities of
interest by substracting the mean and dividing between the standard
deviation of the randomized samples.

@@

\[
\mu= 
\begin{bmatrix}
-0.18\pm 0.04 & -0.41\pm 0.05  -0.20\pm 0.05
\end{bmatrix}
\]

@@

\[
\mu= 
\begin{bmatrix}
-0.17\pm 0.04 & -0.40\pm 0.04 -0.23\pm 0.04\\
\end{bmatrix}
\]

@@

\[
\mu= 
\begin{bmatrix}
-0.32\pm 0.04 -0.74\pm 0.07 -0.08\pm 0.05\\
\end{bmatrix}
\]

>>>>

\begin{table*}
\centering
\begin{tabular}{|p{4.0cm}|p{4.5cm}| p{5.5cm}| c|}\hline
Reference & Target Measurement & Parent Simulation & Probability ($\%$)\\\hline
\text{\cite{2006AJ....131.1405K}} & RMS height in 15 brightest
satellites & Monte Carlo satellite distributions with a power law
radial distribution & $87-99$\\
\text{\cite{2007MNRAS.374.1125M}} & Triaxiality and RMS height in 16 brightest satellites & Monte Carlo from an spherical power law radial
distribution & $17$\\
\text{\cite{2013ApJ...766..120C}}& RMS height in the 27 brightest
satellites & Monte Carlo randomized satellite distribution & High\\
This Work & Triaxiality and RMS height in 11-15 brightest
satellites & 24 halos from selected pairs in a cosmological N-body Dark Matter only simulation ($\sim
10^{6}$\Msun particle mass resolution)& 56 \\
This Work & Triaxiality and RMS height in 11-15 brightest
satellites & 20 halos from selected pairs in a cosmological N-body hydro simulation ($\sim
10^{6}$\Msun particle mass resolution)& 27 \\
This Work & Triaxiality and RMS height in 11-15 brightest
satellites & 12 high resolution DM only N-body simulations ($\sim
10^{5}$\Msun particle mass resolution) of halo pairs & 37 \\
\hline
\end{tabular}
\caption{Probability to find  the triaxiality and/or root mean squared
  (RMS) height of M31 satellites in LCDM simulations. 
\label{table:M31}}
\end{table*}

@@

To summarize, we use asphericity to characterize in the
on equal footing simulations and observations. 
Then, we build an explicit probability distribution for the asphericity and
use simulations to estimate the parameters in this distribution. 
Finally, we use these distributions to generate large number of
samples and directly estimate the number of systems meeting a desired
set of criteria.

@@

\subsection{Describing Samples with the Inertia Tensor}
We base describe the satellites with the inertia
tensor defined by the satellites' positions.

@@

From this tensor we compute its eigenvalues,
$\lambda_1>\lambda_2>\lambda_3$, and corresponding eigenvectors,
$\hat{I}_1$, $\hat{I}_2$, $\hat{I}_3$.
We define the size of the three ellipsoidal axis as
$a=\lambda_1$, $b=\lambda_2$ and $c=\lambda_3$.
We also define $\hat{n}\equiv \hat{I}_1$ as the vector perpendicular to the
planar satellite distribution. 
We also define the Root Mean Squared (RMS) width $w$, of the planar
satellite distribution, as the standard deviation of the satellite
distances to the plane defined by the vector $\hat{n}$.

To summarize we characterize the satellite distribution by for
quantities obtained from the inertia tensor: 
\begin{itemize}
\item RMS plane width, $w$.
\item $c/a$ axis ratio.
\item $b/a$ axis ratio.
\end{itemize}

@@

We compare each satellite distribution against its own spherically
randomized distribution.
We keep fixed the radial position of every satellite
with respect to the central galaxy and then randomize its angular
position. 
We repeat this procedure 1000 times for each satellite distribution
and proceed to measure the quantities mentioned in the previous section:
$w$, $c/a$ and $b/a$.
For each quantity we compute the average and standard deviation from
the 1000 random samples. 
This allows us to build a normalized version of all quantities of
interest by substracting the mean and dividing between the standard
deviation of the randomized samples.

@@

\[
\mu= 
\begin{bmatrix}
-0.18\pm 0.04 & -0.41\pm 0.05 & -0.20\pm 0.05\\
\end{bmatrix}
\]

@@

\[
\mu= 
\begin{bmatrix}
-0.17\pm 0.04 & -0.40\pm 0.04 & -0.23\pm 0.04\\
\end{bmatrix}
\]

@@

\[
\mu= 
\begin{bmatrix}
-0.32\pm 0.04 & -0.74\pm 0.07 &-0.08\pm 0.05\\
\end{bmatrix}
\]
